A Survey on Ordered Binary Decision Diagrams

### 2015-09-15 paper.tex
Revise Abstract and Introduction

\begin{abstract}
This paper serves as a survey on \textit{ordered binary Decision diagrams}
(\textit{OBDDs}). It provides an overview of the representation of OBDDs, the
operations which can be applied efficiently, some aspects when implementing them
as well as their limitations and how to probably overcome them by using
alternative representations. No new results or insights are provided. \tbr
\end{abstract}

@@

In computer-aided design (CAD) as well as other domains like artificial
intelligence or combinatorics, many problems can be formulated as Boolean
functions. These Boolean functions then can be represented symbolically using the
\textit{Binary Decision Diagrams} (\textit{BDDs}) introduced by Lee\cite{LEE59}
and Akers\cite{AKERS78} in 1959 and 1978, respectively.

In 1986 Bryant\cite{BRYANT86} published a paper of strong impact, describing a
\textit{reduced} and \textit{ordered} class of BDDs, so-called \textit{ROBDDs}
(\textit{Reduced Ordered} BDDs) but in literature mostly referred to as
\textit{OBDDs}. OBDDs represent Boolean functions in a \textit{canonical}
form. This results in a compact form and very efficient tests for some properties,
e.g. satisfiability or equivalence\cite{BRYANT86}.

@@

Nonetheless representing Boolean functions using OBDDs has some drawbacks. Classes
of Boolean functions exist for which the size of the graph grows exponentially
independent of the variable ordering, e.g. integer
multiplication~\cite{BRYANT86, BRYANT91, WOELFEL01}.

These limitations of OBDDs motivated researchers to introduce heuristics and
other techniques like \textit{dynamic variable ordering}~\cite{RUDELL93} to solve
the problem of variable ordering~\cite{BOLLIG96}. Furthermore, alternative
representations which e.g. relax the ordering~\cite{BRYANT95} or change the
interpretation of nodes~\cite{BRYANT95, ANDERSEN97} were developed. Such
representations may be beneficial for some classes of Boolean functions but mostly
suffer from losing the canonical form and hence the desirable properties OBDDs
are famous for.

This paper serves as survey on symbolic Boolean function manipulation using OBDDs.
The following section describes the representation of Boolean functions through
BDDs as well as OBDDs. It also highlights the advantages and disadvantages of the
OBDD representation. The third section describes operations which can be applied
to OBDDs efficiently. No source code is provided but those can be looked up in
the corresponding papers easily (references are provided for each algorithm).
Section~\ref{sec:implementational-aspects} describes aspects one has to take into
account when implementing an OBDD package, such as what data structures to use
or how to deal with the problem of variable ordering. Limitations of the OBDD
representation are described in the subsequent section. The last section
discusses alternative representations to possibly overcome the limitations of
the OBDD representation. \tbr

@@

The bit vector are \textit{sparse} in the sense that the set contains much less
than $2^n$ vectors and the bit vectors themselves have many zero elements. ZBDDs
are similar to OBDDs but they use a different reduction rule, exploiting the
sparse bit vectors. In ZBDDs, nodes can be eliminated whenever setting the
corresponding variable to 1 yields 0 as result, which is often the case for sparse
bit vectors. This representation proved successful to solve some combinatorial
problems~\cite{BRYANT95}. \tbr \newpage

@@

\clearpage
% balance columns on last page
%\balance

>>>>

\begin{abstract}
This paper serves as a survey on \textit{ordered binary Decision diagrams}
(\textit{OBDDs}). It provides an overview of the OBDD representation, the
operations which can be applied efficiently on OBDDs and some aspects when it
comes to implementing OBDD packages. Moreover, their limitations and their
application in symbolic model checking are outlined and alternative representations
are presented to tackle these limitations. As this is just a survery, no new
insights are provided.
\end{abstract}

@@

In computer-aided design (CAD) as well as other domains, like artificial
intelligence or combinatorics, many problems can be formulated as Boolean
functions. These Boolean functions then can be represented symbolically using
\textit{Binary Decision Diagrams} (\textit{BDDs}), introduced by Lee~\cite{LEE59}
and Akers~\cite{AKERS78} in 1959 and 1978, respectively.

In 1986, Bryant~\cite{BRYANT86} published a paper having strong impact, describing
a \textit{reduced} and \textit{ordered} class of BDDs, so-called \textit{ROBDDs}
(\textit{Reduced Ordered} BDDs), but in literature mostly referred to as
\textit{OBDDs}. OBDDs represent Boolean functions in a \textit{canonical}
form. This results in a compact form and very efficient tests for some properties,
e.g. satisfiability or equivalence~\cite{BRYANT86}.

@@

Representing Boolean functions using OBDDs has one drawback. Classes of
Boolean functions exist for which the size of the graph grows exponentially
independent of the variable ordering, e.g. integer
multiplication~\cite{BRYANT86, BRYANT91, WOELFEL01}.

These limitations of OBDDs motivated researchers to introduce heuristics and
other techniques like \textit{dynamic variable ordering}~\cite{RUDELL93} to solve
the problem of variable ordering~\cite{BOLLIG96}. Furthermore, alternative
representations were developed which may be beneficial for some classes of Boolean
functions but mostly suffer from losing the canonical form and, hence, the desirable
properties OBDDs are famous for.

This paper serves as survey on symbolic Boolean function manipulation using OBDDs.
The following section describes the representation of Boolean functions through
BDDs as well as OBDDs. It also highlights the advantages and disadvantages of the
OBDD representation. The third section describes operations which can be applied
to OBDDs efficiently. Section~\ref{sec:implementational-aspects} describes aspects
one has to take into account when implementing an OBDD package, such as what data
structures to use or how to deal with the problem of variable ordering.
In the subsequent sections, limitations of the OBDD representation and their usage
in symbolic model checking are outlined. The last section discusses alternative
representations to probably overcome the limitations mentioned before.

@@

The bit vector are \textit{sparse} in the sense that the set contains much less
than $2^n$ vectors and the bit vectors themselves have many zero elements. ZBDDs
are similar to OBDDs but they use a different reduction rule, exploiting the
sparse bit vectors. In ZBDDs, nodes can be eliminated whenever setting the
corresponding variable to 1 yields 0 as result, which is often the case for sparse
bit vectors. This representation proved successful to solve some combinatorial
problems~\cite{BRYANT95}. \tbr %\newpage

@@

%\clearpage
% balance columns on last page
%\balance
